Signatures of Foreground Sky in Power Spectra of Redshifted Neutral Hydrogen from the Epoch of Reionization

### 2014-09-07 foreground_delay_spectrum_modeling.tex
Editing Observation parameters, data figure caption, and primary beam modelling to include discussion about Curtin beam model

One of the primary targets for EoR observing with the MWA is a patch of sky centered around RA = 00h 00m 00.00s, Dec = -30\arcdeg$\,$00\arcmin$\,$00\farcs 0. This is one of the regions of low foreground emission within 27\arcdeg$\,$ from the zenith. The MWA tracks a patch of sky through antenna beams formed and steered electronically by controlling delay settings of an array of dipoles in a MWA tile. These beamformer delay settings can be changed only in discrete steps. The sky is allowed to drift for a certain period of time (usually $\sim 30$~mins) before the discrete shifts in the beamformer delay settings position the beam to be centered again on the same patch of sky. This process is repeated throughout the course of the observation $\approx 4.86$~hours. Figure~\ref{fig:pointings} illustrates how the apparent pointing oscillates around the desired pointing direction.

The LST of two snapshots we chose for our study are shown as vertical lines at -1.92~hours (i.e., 22.08~hours) and 0.09~hours which are hereafter denoted as {\it off--zenith} and {\it zenith} pointings respectively. These pointings (phased array power patterns) are centered at RA = 4\fdg 387, Dec = -29\fdg 86 and RA = -28\fdg 8, Dec = -26\fdg 701, whereas the visibilities themselves are phased to zenith corresponding to RA = -28\fdg 8, Dec = -26\fdg 701, and RA = 1\fdg 35, Dec = -26\fdg 701 respectively.

The observations reported here were observed, near the beginning of the campaign, on Aug 23, 2013. This day has been chosen as a reference data set for comparison within the MWA EoR collaboration. This day was selected on the basis of low number of known system errors from among a long stretch of stable observing.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_fhd_delay_spectrum_snapshots.eps}
\caption{Delay spectrum amplitude, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz), obtained with MWA data for {\it off--zenith} (top) and {\it zenith} (bottom) pointings. White lines mark the boundaries of {\it foreground wedge} determined by the horizon delay limit and antenna spacing. The features in the wedge appear to be brighter and the spillover contamination appears to be higher in the {\it off--zenith} pointing relative to the {\it zenith} pointing due to the presence of brighter foregrounds such as the Galactic center and the Galactic plane. The logarithmic color scale (shown at the right) is common to both panels.\label{fig:fhd_data}}
\end{figure}

@@

\section{Delay Spectrum Modeling}\label{sec:modeling}

We describe power pattern and foreground models we have used in modeling the measured delay spectra.

@@

The power pattern of a MWA tile, $A(\overline{\mathbf{l}},f)$ in equation~\ref{eqn:obsvis}, over the entire hemisphere is analytically modeled based on a 4$\times$4 phased array of isotropic radiators at a height of 0.3~m above an infinite ground plane. We also assume that the delay corrections during the phased addition of dipole voltages in the beamformer suffer from random fluctuations of rms 0.05~ns. The models of power pattern used in data analysis take into account effects of mutual coupling of dipoles in the tile besides finite ground plane effects. Our models were found to be in reasonable agreement with those used in the data analysis.

>>>>

The MWA Epoch of reionization targets two primary low-foreground fields. Here we choose to focus on RA = 0h, Dec = -30\arcdeg.   The MWA tracks a patch of sky through antenna beams formed and steered electronically by controlling delay settings of an array of dipoles in a MWA tile. These beamformer delay settings can be changed only in discrete steps. The sky is allowed to drift for a certain period of time (usually $\sim 30$~mins) before the discrete shifts in the beamformer delay settings position the beam to be centered again on the same patch of sky. This process is repeated throughout the course of the observation $\approx 4.86$~hours. Figure~\ref{fig:pointings} illustrates how the apparent pointing oscillates around the desired pointing direction.

The observations reported here were observed, near the beginning of the campaign, on Aug 23, 2013. This day has been chosen as a reference data set for comparison within the MWA EoR collaboration and was selected on the basis of low number of known system errors from among a long stretch of stable observing.

Here we have chosen two, 112 second long sections from this night for detailed study.  These were chosen to provide a selection of possible foreground and instrumental conditions.  As an example of ``nominal'' observing setup we choose a zenith pointing; as an example of poor foreground conditions we choose a pointing when the field is 2 hours from zenith.  This pointing has a significantly higher secondary side lobe structure and is recorded when the bright galactic center is well above the horizon.

The LSTs of these two snapshots are shown in Figure~\ref{fig:pointings} as vertical lines at -1.92~hours (i.e., 22.08~hours) and 0.09~hours which are hereafter denoted as {\it off--zenith} and {\it zenith} pointings respectively. %These pointings (phased array power patterns) are centered at RA = 4\fdg 387, Dec = -29\fdg 86 and RA = -28\fdg 8, Dec = -26\fdg 701, whereas the visibilities themselves are phased to zenith corresponding to RA = -28\fdg 8, Dec = -26\fdg 701, and RA = 1\fdg 35, Dec = -26\fdg 701 respectively.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_fhd_delay_spectrum_snapshots.eps}
\caption{ Delay spectrum power spectra of two 112s  Murchison Widefield Array observations of the Epoch of Reionization field at RA=0, Dec=-30.  At top an {\it off--zenith}  pointing when tiles have significant side lobe response in the direction of the galactic center. At bottom a {\it zenith} pointing where side lobes are minimal and the galaxy has completely set.  To get the the power spectrum each baseline is Fourier transformed into delay space ($y$-axis) and the amplitude plotted as a logarithmic color in order of baseline length ($x$-axis). The logarithmic color scale (shown at the right) is common to both panels. White lines mark the boundaries of the {\it foreground wedge} determined by the light travel time across each baseline. Smooth spectrum sources are physically limited to the region below this line.  The various features within the wedge depend on the \emph{orientation} of the baseline, the side lobe structure of the primary beam pointing and the sky at that time. \label{fig:fhd_data}}
\end{figure}

@@

\section{Instrument Modeling}\label{sec:modeling}

We describe power pattern and foreground models we have used in modeling the instrument. %todo

@@

The power pattern of a MWA tile ($A(\overline{\mathbf{l}},f)$ in equation~\ref{eqn:obsvis}) is a subject of active research. The most precise model currently available is that of Sutinjo et al (2014; submitted?) which models the 4$\times$4 array of bow-tie dipoles as a mutually-coupled array with the overall power-pattern of the dipole modeled via finite element electromagnetic simulation.  To speed simulation we find that a phased array of isotropic radiators at a height of 0.3~m above an infinite ground plane provides a very good approximation to the detailed simulation. We also assume that the delay corrections during the phased addition of dipole voltages in the beamformer suffer from random fluctuations of rms 0.05~ns, a number in line with the known stability level of the analog signal chain.
